MLH1 Results Outline

### 2020-04-20 doc/MatMeth.Rmd
adding in some calqs to results

**Staining / Immunohistochemistry**

The immunohistochemistry was based on that in [@anderson1999] and [@koehler2002]. To visualize chromosomes in the meiocytes we used markers for the centromere (CREST) and lateral element of of the SC (SYCP3). Double strand breaks (DSB) and COs were visualized with DMC1 and MLH1 respectively. Antibodies and slide blocking was performed in 1X antibody dilution buffer (ADB) (normal donkey serum (Jackson ImmnuoResearch), 1X PBS, bovine serum albumin (Sigma), Triton X-100 (Sigma) ).

@@

SC morphology was used to stage spermatocytes in the zygotene, leptotene, or pachtyene cell stage. Mean **DMC1** foci per cell for a single mouse from a subset of srains *musculus^PWD^*, *musculus^MSM^*, *domesticus^WSB^*, and *domesticus^G^* were compared by t-tests.

### MLH1

@@

@@

X package used to test the fixed effects, Z package used to test the random effects.

@@

>>>>

The **immunohistochemistry** was based on that in [@anderson1999] and [@koehler2002]. To visualize chromosomes in the meiocytes we used markers for the centromere (CREST) and lateral element of of the SC (SYCP3). Double strand breaks (DSB) and COs were visualized with DMC1 and MLH1 respectively. Antibodies and slide blocking was performed in 1X antibody dilution buffer (ADB) (normal donkey serum (Jackson ImmnuoResearch), 1X PBS, bovine serum albumin (Sigma), Triton X-100 (Sigma) ).

@@

Mean **DMC1** foci per cell for a single mouse from a subset of srains *musculus^PWD^*, *musculus^MSM^*, *domesticus^WSB^*, and *domesticus^G^* were compared by t-tests.
SC morphology was used to stage spermatocytes in the zygotene, leptotene, or pachtyene cell stage.

### MLH1 (Analytical Framework)

-Background for setting up/organizing models -- (heiarchal nature of multiple strains within subspecies -- ) - hierarchical nature of categories for inviual mouse averages ()

-(primary focus is mean MLH1 count per cell - at the mouse observation level - apply to more meiocyte traits)

-Our goal -- design an utilize a mixed model for our analytical framework - which leverages / accounts for the hierarchical/nested nature of categories across all the individual mice (e.g. sex, strain subspecies).

@@

**X package used to test the fixed effects, Z package used to test the random effects.**

@@

@@

**combine any X and XY issue with reduced dataset** An additional challenge of the MLH1 framework is that the identities of individual autosomes and the XX in females cannot be easily obtained (the male XY is distinct).

### 2020-04-20 doc/Results.rmd
adding in some calqs to results

<!--
From this figure of the male mouse average of total sc measures, Confirm the general pattern of more SC area for the higher rec strains, compared to the low. But there are some exceptions.

the low rec musc strains which are higher are czech (1) and ast (2) -- the PWD mouse which is lower 24sep14_PWD_m1

INSERT TABLE FOR THE TOTAL SC means ect

The difference between

bin is short for binary, it reports the binarized area of the SC

skel -- takes the skeletonized midline of the bin objects
-->
# Main Figure

@@

@@

>>>>

# Main Figure

@@

**re-read Lenzi et al -- heterogeneity in oogenesis**

@@

**we should note that these interpertations are dependant on which count of DSBs is most reflective for the final number of matur crossovers.** A number of factors make this determintion less clear - including the potential for mulitple rounds of DSB formation (CO homeostasis) (cite) and difference in rates for repair (cite)

(also the data is very good)

### 2020-04-20 doc/Results_v2.rmd
adding in some calqs to results

#this MLH1 only goes up to batch16!
#load(file="C:/Users/alpeterson7/Documents/MLH1repo/data/MLH1/MLH1_data_setup_12.20.19.RData")#batch18
#load(file="C:/Users/alpeterson7/Documents/MLH1repo/data/MLH1/MLH1_data_setup_12.20.19.RData")#fixing bugs, formating
#load(file="data/MLH1/MLH1_data_setup_12.3.19.RData")#removed exclude mice
#load(file="data/MLH1/MLH1_data_setup_11.12.19.RData")#added batch17
#load(file="data/MLH1/MLH1_data_setup_11.11.19.RData")#deleted DUP image
#load(file="data/MLH1/MLH1_data_setup_9.26.19.RData")#added batch16
#load(file="data/MLH1/MLH1_data_setup_8.29.19.RData") #

@@

```{r tot.skel.table, echo=FALSE}
totSC.mouse.av.table <- ddply(.data=MLH1.merge.Skel, 
                 .(mouse),
                 summarize, 
                 ncells = length(unique(Original.Name)),
                 nmice = length(unique(mouse)),
                # mean.MLH1 = mean(adj_nMLH1.foci, na.rm = TRUE),
                # var.MLH1 = var(adj_nMLH1.foci, na.rm = TRUE),

@@

@@

```{r tot.SC.setup, echo = FALSE}

@@

```{r Q2.SC.length.bret.comments, echo=FALSE}
#correlation

@@

@@

Confirming the basic predictions, there is a positive correlation between mean MLH1 foci per cell and total SC **([report spearman's rho and p-value])**. Nevertheless, mean total SC only partially differentiates high-recombination and low-recombination strains (Figure X).  <Bret's note on positive correlation, 'A correlation across mice, across strains, or both?'>

>>>>

#this MLH1 only goes up to batch16!
#load(file="C:/Users/alpeterson7/Documents/MLH1repo/data/MLH1/MLH1_data_setup_12.20.19.RData")#batch18
#load(file="C:/Users/alpeterson7/Documents/MLH1repo/data/MLH1/MLH1_data_setup_12.20.19.RData")#fixing bugs, formating

@@

#adjust 
MLH1.merge.Skel_2$mouse <- as.factor(MLH1.merge.Skel_2$mouse)
MLH1.merge.Skel_2$nMLH1.foci <- as.numeric(MLH1.merge.Skel_2$nMLH1.foci)
MLH1.merge.Skel_2$quality <- as.numeric(MLH1.merge.Skel_2$quality)

```{r tot.skel.table, echo=FALSE}
#calq the correlation for tota.SC and mlh1 - after mouse averages
#use MLH1.merge.Skel_2 -- quality 1-4

totSC.mouse.av.table <- ddply(.data= ( MLH1.merge.Skel_2 %>% filter(quality < 5) ), 
                 .(mouse),
                 summarize, 
                 ncells = length(unique(Original.Name)),
                 nmice = length(unique(mouse)),
                 mean.MLH1 = mean(nMLH1.foci, na.rm = TRUE),
                # var.MLH1 = var(adj_nMLH1.foci, na.rm = TRUE),

@@

#ADD CORRELATION
for(g in length(totSC.mouse.av.table$mouse) ){
totSC.mouse.av.table$cor.SCskel.MLH1[g] <- cor(totSC.mouse.av.table$mean.MLH1[g], totSC.mouse.av.table$mean.skel[g])
}

@@

```{r tot.SC.setup, echo = FALSE}
#calc=q correlation, total.sc and mean MLH1

@@

@@

```{r Q2.SC.length.bret.comments, echo=FALSE}
#correlation, total sc and Mean MLH1

@@

Confirming the basic predictions, there is a positive correlation between mean MLH1 foci per cell and total SC  `r cor(totSC.mouse.av.table$mean.skel, totSC.mouse.av.table$mean.MLH1)`. Nevertheless, mean total SC only partially differentiates high-recombination and low-recombination strains (Figure X).  <Bret's note on positive correlation, 'A correlation across mice, across strains, or both?'>
